Ignition of Fuel Beds by Firebrands

### 2022-05-30 manuscript4.tex
Updates and added model github link

\section{Abstract}
    The increase in wildfire severity and the expansion of the wildland urban interface have increased the need to protect homes from igniting during fires. Firebrand ignition is a significant source of home loss in wildfires. Firebrands often accumulate on or near structures, and an understanding of how multiple firebrands interact with the fuel, environmental factors (i.e., wind), and other firebrands is necessary to better predict ignition and protect homes effectively. This study evaluates changes in the ignition propensity of a fuel bed when two surrogate firebrands are introduced as ignition sources. The fuel beds consisted of Douglas-fir particles between 1.3\si{\milli\meter} and 2.3\si{\milli\meter} in size. Cartridge heaters were used as surrogate firebrands. Four different heater configurations were tested at two different wind conditions. Two heater spacings were used with either two actively heated firebrand surrogates or one inert and one actively heated. A simplified CFD model was implemented to provide additional insights into conditions leading up to ignition. Increasing the wind speed reduced the temperature required for flaming ignition between 20\% and 60\%. Ignition thresholds were largely independent of firebrand surrogate configuration at the high wind speed. However, at the low wind speed, interactions between the firebrand surrogates significantly influenced ignition. Results from the CFD model indicate that ignition occurred in regions where recirculation zones were present. The location of recirculation zones where ignition was initiated was not always located near the firebrand surrogate. More broadly, this work suggests that ...

\section{Introduction}
    The increase of both wildfire severity and the number of homes on the Wildland Urban Interface (WUI) has increased the occurrence of home loss due to wildfires since the turn of the century~\cite{Manzello2013}. Ignition of fuels by firebrands that in turn leads to the loss of a structure is a primary source of home loss~\cite{Koo2010a, Syphard2019, Roberts2021}, and in some fires has been the source of 2/3 of structure ignitions~\cite{Maranghides2013NISTIgnitions}. Thus, it is important to understand the ignition of fuel beds by firebrands in order to mitigate the risk of structure ignition~\cite{Manzello2014}. Ideally, the understanding of the ignition process would result in a predictive model that would enable homeowners, firefighters, and other risk management personnel to determine location specific strategies for preventing ignition before and during a fire. Unfortunately, such a model does not currently exist due to critical shortfalls in the current knowledge of fuel bed ignition. For example, if a burning twig or cone landed on needles accumulated in a gutter of a home there is no model that can quantitatively predict whether or not ignition occurs and therefore the risk of that structure being lost to ignition from the ember cannot be determined. The consequences of  limited risk assesment and mitigation for structures are numerous. For example, in WUI fires firefighting resources are often used to harden structures against firebrand ignition but without an accurate prediction of risk each home must be treated equally which in turn expends fixed resources on homes that don't need assistance at the detriment to those at highest risk.

    Processes that influence ignition begin within the fire when an firebrand (e.g., branch, bark, cone etc) is ignited and lofted into the air by wind. The combusting firebrand is then transported to and lands on a fuel bed near or on a structure. Energy is then transferred to the fuel bed and if the energy is sufficient the fuel bed will undergo pyrolysis and produce flammable gases which may then ignite and begin flaming. These flames may then spread and destroy the structure. Each of these processes, firebrand generation, transport, and ignition, require additional research before a predictive model is possible. This work focuses on the processes that occur during ignition of the fuel bed.

    An additional risk factor for the ignition of structures is the accumulation of firebrands on fuel beds. The geometry of homes often promotes the accumulation of firebrands by creating recirculation zones that promote the deposition of firebrands. The accumulation of firebrands poses an increased risk of home ignition in multiple ways. First, since ignition is largely a stochastic process, the more firebrands that land on an area the larger the probability of ignition. Second, firebrands that accumulate close to one another may depart more energy onto a fuel bed than a single firebrand alone. In work conducted by Hakes et al.~\cite{Hakes2019a} an increase in firebrand pile mass increased the total energy imparted to an instrumented surface. The increase in energy release was attained by a longer duration of energy release as the pile mass increased.

    The accumulation of firebrands near structures also requires the presence of wind which has also been shown to influence ignition. Thus, considering ignition by multiple firebrands also requires the consideration of wind on the process. In work conducted by Suzuki and Manzello~\cite{Suzuki2020a} when wind was increased from 6\si{\meter\per\second} to 8\si{\meter\per\second} the number of firebrands required for ignition of the wood much fuel bed decreased. Similar observations of wind lowering the ignition threshold have been observed in multiple studies with single firebrands. Wang et al.~\cite{Wang2017} reported decreases in ignition times as wind speed increased for hot metal particles dropped on pine needle fuel beds. Ellis reported that the addition of wind increased the threshold of fuel moisture content where ignition was observed for natural firebrands  deposited on eucalypt forest litter. The trends observed by Ellis agreed with those observed by Ganteaume et al.~\cite{Ganteaume2009} and Pulcinski and Anderson~\cite{Plucinski2008} where the addition of wind increased ignition probability. From these studies an increase of wind increases the danger of fuel bed ignition near homes but the mechanism(s) that cause the increased ignition probability are unclear.

    For single firebrands it has been postulated that the primary enhancement of wind is due to increased oxygen to the firebrand and fuel bed. From these works it was unclear if the ignition enhancement was due to increased heat release from the firebrand, increased mixing and oxygen in the pyrolyzates, or some combination of both. Results from the work outlined in Chapter~\ref{chap:manuscript2} show that in the presence of an ignition source where the heat release is not influenced by wind an increase of wind increases the ignition probability. The increase in ignition probability was attributed to the accumulation of pyrolysis products near the energy source (cartridge heater) suggesting that fluid flow near the firebrand is a significant controlling parameter for ignition.

@@

    The combinations of heater spacing and hot or ambient upstream heater were chosen to represent different levels of fluid and thermal interactions between the heaters and the fuel bed. The heater spacing of five diameters was chosen such that minimal interaction between the heaters occurs as previous CFD calculations indicate that the recirculation zone of the upstream heater is approximately five diameters when the heater is in a wind of 5.8 \si{\meter\per\second} and an orientation perpendicular to the flow. Preliminary thermal calculations also indicated that the pyrolysis fronts created by each heater are unlikely to interact within previously observed times to ignition. The heater spacing of one diameter was chosen for a high level of interaction between both the fluid disturbances and thermal fronts of each firebrand. The one diameter spacing places the downstream heater inside the recirculation zone of the upstream heater under wind and the preliminary heat transfer calculations indicated that the thermal fronts of each heater will merge before the anticipated time to ignition. The chosen heater orientations also represent potential scenarios that may be encountered in a wildfire. The configuration with two heaters both heated is representative of a multi firebrand attack where the one diameter spacing approximates a firebrand pile and the five diameter spacing approximates two firebrands falling in close proximity but not within each others region of influence. The configuration with only the downstream heater heated is representative of an firebrand falling near an object (e.g., twig, rock, or cone) that disturbs the flow and may act as a heat sink for the energy deposited to the firebrand with the one and five diameter spacing representing an firebrand falling both within and outside of the region of influence of the firebrand. The configuration where the upstream heater is not heated and five diameters is also considered a control for comparison to previous single heater ignition results.

@@

    The cartridge heaters used had diameters of 6.4\si{\milli\meter} and lengths of 51\si{\milli\meter}. The heater sizes were chosen to represent large firebrands with a high potential to ignite a fuel bed in a wildfire~\cite{Manzello2007}. The heater temperatures were controlled using a PID temperature controller implemented in LabVIEW. Heater temperatures ranged from 250\si{\celsius} to 750\si{\celsius}. Heater temperatures were measured using a type-K thermocouple attached to the center of each heater opposite the fuel bed.
    Controlling the temperature of the heater provides an advantage over natural burning or smoldering firebrands and pre-heated particles by removing the temperature variability of the heat source and enabling real time data logging of the heat source temperature. Controlling the temperature of the firebrand also removes some complexity of calculating the energy imparted to the fuel bed. The heater was lowered onto the fuel bed with an automated lowering device. The lowering device included a load cell and a PID controller to maintain a force equivalent to a two 10\si{\gram} firebrands throughout the experiment. To minimize flow disruptions due to apparatus the heaters were each attached to two 4-40 threaded rods that extended approximately 100\si{\milli\meter} below the load cell.

@@

    The fuel bed material was processed from kiln dried Douglas-fir lumber. The lumber was first planed to generate shavings. The wood shavings were then granulated and screened such that the particles passed through a 2.3\si{\milli\meter} screen but not through a 1.3\si{\milli\meter} screen. The particles were then dried in an oven at 103\si{\celsius} to remove any remaining moisture content. During tests, the fuel was placed in a 140\si{\milli\meter} diameter glass container with a depth of 70\si{\milli\meter} and then inserted into the wind tunnel. The average density of the fuel beds was 93\si{\kilo\gram\per\cubic\meter}.

    A simplified model was implemented to obtain further insight into the fluid mechanics and heat transfer processes that influence the trends observed in the experimental efforts. The model was implemented in three parts. First, a thermal model of the fuel bed was used to estimate the mass of the fuel bed above the pyrolysis temperature and the average temperature. Second, the mass of the fuel bed material converted from solid fuel to gaseous pyrolysis products was estimated. In the third and final step the fluid flow and heat transfer between the heaters, air, and released pyrolyzates was modeled. The modelling efforts were conducted for a 10\si{\second} interval to ensure proper characterization of the flow field and maintain an accurate representation of the fuel bed. As mass is lost from the fuel bed due to pyrolysis the contact area between the heater and the fuel bed changes as does the surface shape of the fuel bed. It was observed during experiments that the contact between the heat and the fuel bed and the shape of the fuel bed itself began to deviate significantly from the model domain at approximately 10\si{\second}. Considerations of changes in contact area and changing geometry due to pyrolysis is out of the scope of this work. Nonetheless, the treatment of each configuration equally and within the limited time frame is still anticipated to produce insight into the influence of the thermal and fluid processes on ignition.

    The thermal model of the fuel bed was conducted using OpenFOAM~\cite{Foundation2020}. Figure~\ref{fig:multiHeaterThermal} shows the computational domain of the fuel bed. The thermal diffusivity of the fuel bed materials was estimated using correlations from literature for the thermal conductivity~\cite{bergman2011fundamentals} and published values of specific heat~\cite{Parker1986}. Equation~\ref{eqn:estk4} shows the the correlation for effective thermal conductivity where $\epsilon$, as defined in Equation~\ref{eqn:epsilon4}, is the proportion of the fuel bed that is air. The bulk density of the fuel bed material ($\rho_{bed}$) and the solid wood ($\rho_{solid}$) were obtained from experimental samples. 
        \begin{equation}
            k_{eff} = \frac{1}{2} \left(\frac{1}{\left( 1 - \epsilon \right)/k_{solid} + \epsilon/k_{air}} +  \epsilon k_{air} + \left(1-\epsilon \right) k_{solid} \right)
            \label{eqn:estk4}
        \end{equation}
        \begin{equation}
            \epsilon = 1- \frac{\rho_{solid}}{\rho_{bed}}
            \label{eqn:epsilon4}
        \end{equation}
    The interface between the fuel bed and air and the radial edge of the domain were treated as insulated. The centerline was treated as a symmetric boundary condition. The insulated radial edge and symmetric boundary condition are consistent with boundary conditions in the actual experiment since the 25\si{\milli\meter} distance ensures no temperature change during 10\si{\second} of heating modeled and the heater temperature has been observed to be circumferentially consistent with infrared imaging. The insulated interface with the air does not account for radiation and convective heat losses from the surface of the fuel bed as it is heated. However, the magnitude of the heat losses from the fuel bed is estimated to be significantly smaller in magnitude than the heat transfer from the heater. Reactions, mass loss due to pyrolysis, and changes in contact area due to changing fuel bed geometry are not considered in the temperature analysis. While the aforementioned assumptions and limitations reduce the accuracy of the model in comparison to the experimental conditions, the temperature distributions still provide insights into differences in the mass of the fuel bed material that undergoes pyrolysis and provides insight to better interpret the experimental results. 
    \begin{figure}
        \centering
        \resizebox{0.375\textwidth}{!}{%
          \begin{tikzpicture}
            \node at (0,0) (origin) {};
            %\draw (origin) rectangle (topright);
            \fill[pattern=north west lines, pattern color= brown, draw=black] (3.18mm, 0) -- (25mm, 0) arc (0:-90:25mm) -- (0, -3.18mm) arc (270:0:3.18mm ) -- cycle ;

@@

            \fill[pattern=north east lines, pattern color=black!90] (-2mm,-3.18mm) rectangle (0mm,-27mm) node[rotate=-90, pos=0.5] {};
            %\fill[pattern=north east lines, pattern color=black!90] (0mm,-2mm) rectangle (27mm,0mm) node[rotate=-90, pos=0.5] {};
           \fill[pattern=crosshatch] (25mm, 0) -- (27mm, 0) arc (0:-90:27mm) -- (0, -25mm) arc(-90:0:25mm) ;
            \draw[-stealth,decorate,decoration={snake,amplitude=1pt,pre length=0pt,post length=2pt}] (3.5mm, -6mm) -- ++(1mm, 9mm);
            \draw[-stealth,decorate,decoration={snake,amplitude=1pt,pre length=0pt,post length=2pt}] (6mm, -5mm) -- ++(1mm, 7mm);
            \draw [|-|] (-5mm, 0mm) -- (-5mm, -25mm) node[pos=0.5, left] {25mm} ;
             \draw[thick, dashed] (3.18mm, 0) -- (10mm, 0);
            \draw (0mm, -3.18mm) -- (0mm, -10mm);
            \node[draw, rectangle, fill=white, draw=white] (p) at (8mm, 10mm) {Pyrolysis Gases};
	        \draw[->, thick] (p.south) -- (6mm, 2mm);
	        \node[draw, rectangle, fill=white, draw=white, anchor=west] (p) at (12mm, 4mm) {T=220\si{\celsius}};
	        \draw[->, thick] (p.west) -- (10mm, 1mm);
          \end{tikzpicture}
          }
        \caption{Diagram of the computational domain where black lines indicate domain boundaries, and red lines are boundaries defined by the heater. The arrows denote flow of pyrolysis products from the fuel bed into the air above the fuel bed.}
        \label{fig:multiHeaterThermal}
    \end{figure}

    The velocity of pyrolyzates entering the fluid domain from the fuel bed was determined in a two step process. First the mass of material in the fuel bed undergoing pyrolysis was estimated. All material over a temperature of 220\si{\celsius} was considered to undergo pyrolysis. This temperature was chosen as 220\si{\celsius} corresponds to the minimum temperature for the pyrolysis of hemicellulose and is considered the onset of pyrolysis~\cite{Yang2007a}. The amount of mass estimated to depart the fuel bed was determined to be the proportion of fuel bed material that converted to gas phase products when reacted at the average temperature of the material above the pyrolysis threshold from the thermal model. Reactions were conducted using the BioPOx mechanism~\cite{Dhahak2019} and mass was assumed to depart the fuel bed if it existed in the gas phase. Species were considered to be gas phase if they were present in the Bio1412 mechanism~\cite{Ranzi2001, Ranzi2008}, since the Bio1412 mechanism considers the gas phase reaction of pyrolysis products. The reaction calculations were conducted in Cantera~\cite{Goodwin2020}. The Cantera calculations were 0D which necessitated using the average temperatures of the fuel bed for the pyrolysis calculations rather than spatially resolved temperatures. Mass flux of the gases exiting the fuel bed was determined by averaging the total mass of gaseous pyrolysis products produced over the 10\si{\second} interval. The exit velocity was then determined as the mass flux through the distance between the the heater edge and the pyrolysis threshold as shown by the dotted line in Figure~\ref{fig:multiHeaterThermal}. The calculated velocity and injection area was then used as an input for calculations conducted in the flow field.

    A 2D RANS model of the flow above the fuel bed and around the heaters was implemented in OpenFOAM. The area of the computational domain with respect to the experimental apparatus is shown as the dashed line in Figure~\ref{fig:multiHeaterApparatus}. The 2D slice of the domain modeled was considered to pass through the center of the heaters and corresponds to the location of the highest temperature of the heater which also typically coincides with the ignition location. Turbulence parameters were determined from measurements taken of the experimental apparatus with a TSI IFA-300 hot wire anemometer. The k-$\epsilon$ turbulence model was used with values of 1.4$\cdot$10\textsuperscript{-4}\si{\square\meter\per\square\second} and 9.5$\cdot$10\textsuperscript{-6}\si{\square\meter\per\cubic\second} for k and $\epsilon$ respectively. Calculations were conducted for 10\si{\second} and were then averaged to observed differences in temperature and velocity characteristics between the different configurations.

>>>>

\section{Abstract}
    The increase in wildfire severity and the expansion of the wildland urban interface have increased the need to protect homes from igniting during fires. Firebrand ignition is a significant source of home loss in wildfires. Firebrands often accumulate on or near structures, and an understanding of how multiple firebrands interact with the fuel, environmental factors (i.e., wind), and other firebrands is necessary to better predict ignition and protect homes effectively. This study evaluates changes in the ignition propensity of a fuel bed when two surrogate firebrands are introduced as ignition sources. The fuel beds consisted of Douglas-fir particles between 1.3\si{\milli\meter} and 2.3\si{\milli\meter} in size. Cartridge heaters were used as surrogate firebrands. Four different heater configurations were tested at two different wind conditions. Two heater spacings were used with either two actively heated firebrand surrogates or one inert and one actively heated. A simplified CFD model was implemented to provide additional insights into conditions leading up to ignition. Increasing the wind speed reduced the temperature required for flaming ignition between 20\% and 60\%. Ignition thresholds were largely independent of firebrand surrogate configuration at the high wind speed. However, at the low wind speed, interactions between the firebrand surrogates significantly influenced ignition. Results from the CFD model indicate that ignition occurred in regions where recirculation zones were present. The location of recirculation zones where ignition was initiated was not always located near the firebrand surrogate. More broadly, this work suggests that at low wind speeds or in quiescent conditions, thermal interactions between firebrands has a significant impact on ignition, but higher wind speeds, fluid dynamic effects control ignition.

\section{Introduction}
    The increase of both wildfire severity and the number of homes on the Wildland Urban Interface (WUI) has increased the occurrence of home loss due to wildfires since the turn of the century~\cite{Manzello2013}. Ignition of fuels by firebrands that in turn leads to the loss of a structure is a primary source of home loss~\cite{Koo2010a, Syphard2019, Roberts2021}, and in some fires, has been the source of 2/3 of structure ignitions~\cite{Maranghides2013NISTIgnitions}. Thus, it is important to understand the ignition of fuel beds by firebrands in order to mitigate the risk of structure ignition~\cite{Manzello2014}. Ideally, the understanding of the ignition process would result in a predictive model that would enable homeowners, firefighters, and other risk management personnel to determine location-specific strategies for preventing ignition before and during a fire. Unfortunately, such a model does not currently exist due to critical shortfalls in the current knowledge of fuel bed ignition. For example, if a burning twig or cone landed on needles accumulated in a gutter of a home, there is no model that can quantitatively predict whether or not ignition occurs, and therefore the risk of that structure being lost to ignition from the ember cannot be determined. The consequences of limited risk assessment and mitigation for structures are numerous. For example, in WUI fires, firefighting resources are often used to harden structures against firebrand ignition, but without an accurate prediction of risk, each home must be treated equally, which in turn expends fixed resources on homes that don't need assistance at the detriment to those at highest risk.

    Processes that influence ignition begin within the fire when a firebrand (e.g., branch, bark, cone, etc.) is ignited and lofted into the air by wind. The combusting firebrand is then transported to and lands on a fuel bed near or on a structure. Energy is then transferred to the fuel bed, and if the energy is sufficient, the fuel bed will undergo pyrolysis and produce flammable gases, which may then ignite and begin flaming. These flames may then spread and destroy the structure. Each of these processes, firebrand generation, transport, and ignition, require additional research before a predictive model is possible. This work focuses on the processes that occur during ignition of the fuel bed.

    An additional risk factor for the ignition of structures is the accumulation of firebrands on fuel beds. The geometry of homes often promotes the accumulation of firebrands by creating recirculation zones that promote the deposition of firebrands. The accumulation of firebrands poses an increased risk of home ignition in multiple ways. First, since ignition is largely a stochastic process, the more firebrands that land on an area, the larger the probability of ignition. Second, firebrands that accumulate close to one another may depart more energy onto a fuel bed than a single firebrand alone. In work conducted by Hakes et al.~\cite{Hakes2019a} an increase in firebrand pile mass increased the total energy imparted to an instrumented surface. The increase in energy release was attained by a longer duration of energy release as the pile mass increased.

    The accumulation of firebrands near structures also requires the presence of wind which has also been shown to influence ignition. Thus, considering ignition by multiple firebrands also requires the consideration of wind on the process. In work conducted by Suzuki and Manzello~\cite{Suzuki2020a} when wind was increased from 6\si{\meter\per\second} to 8\si{\meter\per\second} the number of firebrands required for ignition of the wood much fuel bed decreased. Similar observations of wind lowering the ignition threshold have been observed in multiple studies with single firebrands. Wang et al.~\cite{Wang2017} reported decreases in ignition times as wind speed increased for hot metal particles dropped on pine needle fuel beds. Ellis reported that the addition of wind increased the threshold of fuel moisture content where ignition was observed for natural firebrands deposited on eucalypt forest litter. The trends observed by Ellis agreed with those observed by Ganteaume et al.~\cite{Ganteaume2009} and Pulcinski and Anderson~\cite{Plucinski2008} where the addition of wind increased ignition probability. From these studies, an increase in wind increases the danger of fuel bed ignition near homes, but the mechanism(s) that cause the increased ignition probability are unclear.

    For single firebrands, it has been postulated that the primary enhancement of wind is due to increased oxygen to the firebrand and fuel bed. From these works it was unclear if the ignition enhancement was due to increased heat release from the firebrand, increased mixing and oxygen in the pyrolyzates, or some combination of both. Results from the work outlined in Chapter~\ref{chap:manuscript2} show that in the presence of an ignition source where the heat release is not influenced by wind an increase of wind increases the ignition probability. The increase in ignition probability was attributed to the accumulation of pyrolysis products near the energy source (cartridge heater) suggesting that fluid flow near the firebrand is a significant controlling parameter for ignition.

@@

    The combinations of heater spacing and hot or ambient upstream heater were chosen to represent different levels of fluid and thermal interactions between the heaters and the fuel bed. The heater spacing of five diameters was chosen such that minimal interaction between the heaters occurs as preliminary CFD calculations indicated that the recirculation zone of the upstream heater is approximately five diameters when the heater is in a wind of 5.8 \si{\meter\per\second} and an orientation perpendicular to the flow. Preliminary thermal calculations also indicated that the pyrolysis fronts created by each heater are unlikely to interact within previously observed times to ignition. The heater spacing of one diameter was chosen for a high level of interaction between both the fluid disturbances and thermal fronts of each firebrand. The one diameter spacing places the downstream heater inside the recirculation zone of the upstream heater under wind, and the preliminary heat transfer calculations indicated that the thermal fronts of each heater will merge before the anticipated time to ignition.

    The chosen heater orientations also represent potential scenarios that may be encountered in a wildfire. The configuration with two heaters both heated is representative of a multi firebrand attack where the one diameter spacing approximates a firebrand pile and the five diameter spacing approximates two firebrands falling in close proximity but not within each others region of influence. The configuration with only the downstream heater heated is representative of a firebrand falling near an object (e.g., twig, rock, or cone) that disturbs the flow and may act as a heat sink for the energy deposited to the firebrand with the one and five diameter spacing representing a firebrand falling both within and outside of the region of influence of the firebrand. The configuration where the upstream heater is not heated and five diameters is also considered a control for comparison to previous single heater ignition results.

@@

    The cartridge heaters used had diameters of 6.4\si{\milli\meter} and lengths of 51\si{\milli\meter}. The heater sizes were chosen to represent large firebrands with a high potential to ignite a fuel bed in a wildfire~\cite{Manzello2007}. The heater temperatures were controlled using a PID temperature controller implemented in LabVIEW. Heater temperatures ranged from 250\si{\celsius} to 750\si{\celsius}. Heater temperatures were measured using a type-K thermocouple attached to the center of each heater opposite the fuel bed.
    Controlling the temperature of the heater provides an advantage over natural burning or smoldering firebrands and pre-heated particles by removing the temperature variability of the heat source and enabling real-time data logging of the heat source temperature. Controlling the temperature of the firebrand also removes some complexity of calculating the energy imparted to the fuel bed. The heater was lowered onto the fuel bed with an automated lowering device. The lowering device included a load cell and a PID controller to maintain a force equivalent to a two 10\si{\gram} firebrands throughout the experiment. To minimize flow disruptions due to apparatus, the heaters were each attached to two 4-40 threaded rods that extended approximately 100\si{\milli\meter} below the load cell.

@@

    The fuel bed material was processed from kiln-dried Douglas-fir lumber. The lumber was first planed to generate shavings. The wood shavings were then granulated and screened such that the particles passed through a 2.3\si{\milli\meter} screen but not through a 1.3\si{\milli\meter} screen. The particles were then dried in an oven at 103\si{\celsius} to remove any remaining moisture content. During tests, the fuel was placed in a 140\si{\milli\meter} diameter glass container with a depth of 70\si{\milli\meter} and then inserted into the wind tunnel. The average density of the fuel beds was 93\si{\kilo\gram\per\cubic\meter}.

    A simplified model was implemented to obtain further insight into the fluid mechanics and heat transfer processes that influence the trends observed in the experimental efforts. The model was implemented in three parts. First, a thermal model of the fuel bed was used to estimate the mass of the fuel bed above the pyrolysis temperature and the average temperature. Second, the mass of the fuel bed material converted from solid fuel to gaseous pyrolysis products was estimated. In the third and final step, the fluid flow and heat transfer between the heaters, air, and released pyrolyzates was modeled. The modeling efforts were conducted for a 10\si{\second} interval to ensure proper characterization of the flow field and maintain an accurate representation of the fuel bed. As mass is lost from the fuel bed due to pyrolysis, the contact area between the heater and the fuel bed changes as does the surface shape of the fuel bed. It was observed during experiments that the contact between the heat and the fuel bed and the shape of the fuel bed itself began to deviate significantly from the model domain at approximately 10\si{\second}. Considerations of changes in contact area and changing geometry due to pyrolysis is out of the scope of this work. Nonetheless, the treatment of each configuration equally and within the limited time frame is still anticipated to produce insight into the influence of the thermal and fluid processes on ignition.

    The thermal model of the fuel bed was conducted using OpenFOAM~\cite{Foundation2020}. Figure~\ref{fig:multiHeaterThermal} shows the computational domain of the fuel bed. The thermal diffusivity of the fuel bed materials was estimated using correlations from literature for the thermal conductivity~\cite{bergman2011fundamentals} and published values of specific heat~\cite{Parker1986}. Equation~\ref{eqn:estk4} shows the correlation for effective thermal conductivity where $\epsilon$, as defined in Equation~\ref{eqn:epsilon4}, is the proportion of the fuel bed that is air. The bulk density of the fuel bed material ($\rho_{bed}$) and the solid wood ($\rho_{solid}$) were obtained from experimental samples. 
        \begin{equation}
            k_{eff} = \frac{1}{2} \left(\frac{1}{\left( 1 - \epsilon \right)/k_{solid} + \epsilon/k_{air}} +  \epsilon k_{air} + \left(1-\epsilon \right) k_{solid} \right)
            \label{eqn:estk4}
        \end{equation}
        \begin{equation}
            \epsilon = 1- \frac{\rho_{solid}}{\rho_{bed}}
            \label{eqn:epsilon4}
        \end{equation}
    The interface between the fuel bed and air and the radial edge of the domain were treated as insulated. The centerline was treated as a symmetric boundary condition. The insulated radial edge and symmetric boundary conditions are consistent with boundary conditions in the actual experiment since the 25\si{\milli\meter} distance ensure no temperature change during 10\si{\second} of heating modeled, and the heater temperature has been observed to be circumferentially consistent with infrared imaging. The insulated interface with the air does not account for radiation and convective heat losses from the surface of the fuel bed as it is heated. However, the magnitude of the heat losses from the fuel bed is estimated to be significantly smaller in magnitude than the heat transfer from the heater. Reactions, mass loss due to pyrolysis, and changes in contact area due to changing fuel bed geometry are not considered in the temperature analysis. While the aforementioned assumptions and limitations reduce the accuracy of the model in comparison to the experimental conditions, the temperature distributions still provide insights into differences in the mass of the fuel bed material that undergoes pyrolysis and provides insight to better interpret the experimental results. 
    \begin{figure}
        \centering
        \resizebox{0.375\textwidth}{!}{%
          \begin{tikzpicture}
            \node at (0,0) (origin) {};
            %\draw (origin) rectangle (topright);
            \fill[pattern=north west lines, pattern color= brown, draw=black] (3.18mm, 0) -- (25mm, 0) arc (0:-90:25mm) -- (0, -3.18mm) arc (270:0:3.18mm ) -- cycle ;

@@

            \fill[pattern=north east lines, pattern color=black!90] (-2mm,-3.18mm) rectangle (0mm,-27mm) node[rotate=-90, pos=0.5] {};
            %\fill[pattern=north east lines, pattern color=black!90] (0mm,-2mm) rectangle (27mm,0mm) node[rotate=-90, pos=0.5] {};
           \fill[pattern=crosshatch] (25mm, 0) -- (27mm, 0) arc (0:-90:27mm) -- (0, -25mm) arc(-90:0:25mm) ;
            \draw[-stealth,decorate,decoration={snake,amplitude=1pt,pre length=0pt,post length=2pt}] (3.5mm, -6mm) -- ++(1mm, 9mm);
            \draw[-stealth,decorate,decoration={snake,amplitude=1pt,pre length=0pt,post length=2pt}] (6mm, -5mm) -- ++(1mm, 7mm);
            \draw [|-|] (-5mm, 0mm) -- (-5mm, -25mm) node[pos=0.5, left] {25mm} ;
             \draw[thick, dashed] (3.18mm, 0) -- (10mm, 0);
            \draw (0mm, -3.18mm) -- (0mm, -10mm);
            \node[draw, rectangle, fill=white, draw=white] (p) at (8mm, 10mm) {Pyrolysis Gases};
	        \draw[->, thick] (p.south) -- (6mm, 2mm);
	        \node[draw, rectangle, fill=white, draw=white, anchor=west] (p) at (12mm, 4mm) {T=220\si{\celsius}};
	        \draw[->, thick] (p.west) -- (10mm, 1mm);
          \end{tikzpicture}
          }
        \caption{Diagram of the computational domain where black lines indicate domain boundaries and red lines are boundaries defined by the heater. The arrows denote flow of pyrolysis products from the fuel bed into the air above the fuel bed.}
        \label{fig:multiHeaterThermal}
    \end{figure}

    The velocity of pyrolyzates entering the fluid domain from the fuel bed was determined in a two-step process. First, the mass of material in the fuel bed undergoing pyrolysis was estimated. All material over a temperature of 220\si{\celsius} was considered to undergo pyrolysis. This temperature was chosen as 220\si{\celsius} corresponds to the minimum temperature for the pyrolysis of hemicellulose and is considered the onset of pyrolysis~\cite{Yang2007a}. The amount of mass estimated to depart the fuel bed was determined to be the proportion of fuel bed material that converted to gas-phase products when reacted at the average temperature of the material above the pyrolysis threshold from the thermal model. Reactions were conducted using the BioPOx mechanism~\cite{Dhahak2019}, and mass was assumed to depart the fuel bed if it existed in the gas phase. Species were considered to be gas-phase if they were present in the Bio1412 mechanism~\cite{Ranzi2001, Ranzi2008} since the Bio1412 mechanism considers the gas-phase reaction of pyrolysis products. The reaction calculations were conducted in Cantera~\cite{Goodwin2020}. The Cantera calculations were 0D which necessitated using the average temperatures of the fuel bed for the pyrolysis calculations rather than spatially resolved temperatures. Mass flux of the gases exiting the fuel bed was determined by averaging the total mass of gaseous pyrolysis products produced over the 10\si{\second} interval. The exit velocity was then determined as the mass flux through the distance between the heater edge and the pyrolysis threshold, as shown by the dotted line in Figure~\ref{fig:multiHeaterThermal}. The calculated velocity and injection area was then used as an input for calculations conducted in the flow field.

    A 2D RANS model of the flow above the fuel bed and around the heaters was implemented in OpenFOAM. The area of the computational domain with respect to the experimental apparatus is shown as the dashed line in Figure~\ref{fig:multiHeaterApparatus}. The 2D slice of the domain modeled was considered to pass through the center of the heaters and corresponds to the location of the highest temperature of the heater, which also typically coincides with the ignition location. Turbulence parameters were determined from measurements taken of the experimental apparatus with a TSI IFA-300 hot wire anemometer. The k-$\epsilon$ turbulence model was used with values of 1.4$\cdot$10\textsuperscript{-4}\si{\square\meter\per\square\second} and 9.5$\cdot$10\textsuperscript{-6}\si{\square\meter\per\cubic\second} for k and $\epsilon$ respectively. Calculations were conducted for 10\si{\second} and were then averaged to observe differences in temperature and velocity characteristics between the different configurations.

\section{Results and Discussion}
    The flaming ignition or non-ignition result of each test is shown in Figure~\ref{fig:multi_heater_hot_vs_ambient}. The markers show the result of each test, and the curves show the logistic regression for each of the heater configurations as defined in Table~\ref{tab:multiHeaterConfig}. The shaded regions around the curves represent the 95\% confidence intervals of each regression. The top plot in Figure~\ref{fig:multi_heater_hot_vs_ambient} shows configurations where the downstream heater is actively heated and the upstream heater is at ambient conditions. The bottom plot in Figure~\ref{fig:multi_heater_hot_vs_ambient} shows results for configurations where both heaters are actively heated to the same temperature.
        \begin{figure}[hpbt]
            \centering
            \includegraphics[width=0.75\columnwidth]{Figures/multi_heater_plot.png}
            \caption{Ignition or no ignition outcomes of tests with different heater configurations. The circular markers indicate outcomes of individual tests and the curves represent logistic regressions of each test series. The shaded region shows the 95\% confidence interval for each regression.}
            \label{fig:multi_heater_hot_vs_ambient}
        \end{figure}
    Table~\ref{tab:multiFiftyTemp} shows the heater temperatures estimated to result in a 50\% ignition probability from the logistic regressions shown in Figure~\ref{fig:multi_heater_hot_vs_ambient} as well  Temperature values for the corresponding wind speed and orientation from the single heater study in Chapter~\ref{chap:manuscript2} are shown in the column titled "Single" for comparison. 
        \rowcolors{2}{gray!25}{white}
        \begin{table}[hpbt]
            \normalsize
            \caption{Heater temperature required to achieve 50\% ignition probability for the configurations tested with 95\% confidence intervals shown in parentheses. The column "Unheated" refers to cases where the upstream heater was not temperature controlled. Hence the temperature corresponds to the temperature of the hot heater. "Heated" refers to cases where both upstream and downstream heaters were maintained at the setpoint temperature. The single heater column shows predicted ignition temperatures from the study in Chapter~\ref{chap:manuscript2} for a single heater.}
            \centering
            \begin{tabular}{ccrrr}
                \rowcolor{gray!50}
               U\textsubscript{bulk} (\si{\meter\per\second}) & Spacing (D) & Unheated (\si{\celsius})& Heated (\si{\celsius}) & Single (\si{\celsius})\\
                \hline
                0.5  & 1 & 750 (644, 850) & 533 (527, 541) & 571\\
                0.5  & 5 & 578 (565, 583) & 601 (597, 604) & 571\\
                5.8  & 1 & 423 (417, 429) & 406 (397, 416) & 399\\
                5.8  & 5 & 438 (421, 454) & 396 (387, 407) & 399
            \end{tabular}
            \label{tab:multiFiftyTemp}
        \end{table}
